A Study of Common Bug Fix Patterns in Rust

### 2022-11-27 discussion.tex
discussion

In this work, using Ruxanne we successfully mined 20 cross project bug fix patterns. We introduced these patterns in two groups: 12 general patterns and 8 BC-related patterns. Clippy is able to detect five of the BC-related patterns. Fundamentally, Clippy cannot detect some of them. For instance, we proposed the pattern \textit{Adding mutability} which Clippy cannot detect, simply because it cannot predict the next change that the developer is going to make. To the best of our knowledge, the rest of the patterns cannot be reported by the current linting tools. We therefore believe that our work introduces a new direction for IDE tools to help Rust developers.

In each group, we categorized the bug fix patterns based on the underlying program element. The patterns encompass a wide variety of program elements both in the general and BC-related patterns. This shows that our weighting scheme is not biased towards specific program elements. Also, our patterns are all cross-project fix patterns: each fix pattern has been seen at least in three different projects.

Furthermore, we expected to obtain some patterns associated with certain Rust-related features but our results did not include those patterns. Our speculation is that, since Rust has a strong compiler check before creating the program executable, the programmers usually get many errors, and they only push their code once it is free of compile errors. For instnace, we did not observe any patterns associated with \verb+move+ operator in Rust. This operator is used to transfer the ownership of values to a closure. In some cases (e.g. multithreading) not moving the ownership can create invalid references, something that Rust's compiler prevents. That is why we believe due to Rust's strong compiler check some patterns were hidden from our sight.

We believe our results can help researchers to create IDE tools for Rust. For instance, we discussed that \textit{Modifying the attributes of structs} is a common pattern that is observed in bug fix code changes. Assuming that the correct usage of attributes can be found other locations in the codebase~\citep{forrest2009genetic}, a code linter tool can be crafted to look for correct attribute list.

Similarly, we presented the pattern \textit{Dropping clone and adding borrowing}. As discussed, Clippy does not detect this pattern, and repetitive clones can be computationally expensive and might lead to hogging the CPU. Using state-of-the-art method for program repair, a tool can be designed to recognize this pattern and change variable cloning to a simple borrowing.

A general approach for code embedding is to extract paths with semantic information that serves better to our embedding goal. For instance, to design an embedding method for repairing Rust programs, the designer might want to disregard certain elements, such as moving ownerships to closure body---as they did not appear in the bug patterns we presented.

All of our parsing, path extraction, weighting scheme and clustering modules in addition to our final results can be found in our replication package for verification, resuability and further extension. The weighting can be modified to make the embedding focus on a specific set of elements for building different code embeddings.

>>>>

In this work, we used Ruxanne to successfully mine 20 cross-project bug fix patterns. We presented these patterns in two groups: 12 general patterns and 8 BC-related patterns. Clippy is able to detect five of the BC-related patterns. Fundamentally, Clippy cannot detect some of the remaining patterns. For instance, we proposed the pattern \textit{Adding mutability}. In the Rust context, this change anticipates a future change to the code which will perform a mutation of a previously immutable value. Predicting the future is beyond Clippy's remit. To the best of our knowledge, the rest of the patterns cannot be reported by the current linting tools.

In each group, we categorized the bug fix patterns based on their underlying program element. The patterns encompass a wide variety of program elements both in the general and BC-related groups. This shows that our weighting scheme is not biased towards specific program elements. Also, our patterns are all cross-project fix patterns: each fix pattern has been seen at least in three different projects.

Furthermore, we expected to obtain some patterns associated with certain Rust-related features. However, our results did not include those patterns. For instance, we did not observe any patterns associated with the Rust \verb+move+ operator, which has been known to trouble many beginning Rust programmers. \verb+move+ is used to transfer the ownership of values to a closure. In some cases (e.g. multithreading), not moving the ownership could create invalid references. Rust compiler checks prevent such errors. Our speculation is that, since Rust enforces strict compiler checks, programmers usually get many errors, and they only push their code once it is free of compile errors. We thus would not observe patterns which don't make it to the repository.

We believe our results can help researchers to create IDE tools for Rust. For instance, we discussed that \textit{Modifying the attributes of structs} is a common pattern that is observed in bug fix code changes. Assuming that the correct usage of attributes can be found other locations in the codebase~\citep{forrest2009genetic}, a code linter tool can be crafted to look for correct attribute lists.

Similarly, we presented the pattern \textit{Dropping clone and adding borrowing}. As discussed, Clippy does not detect this pattern, and repetitive clones can be computationally expensive and might lead to hogging the CPU. Using state-of-the-art methods for program repair, one could design a tool to recognize this pattern and change variable cloning to simple borrowing.

A general approach for code embedding is to extract paths with semantic information that better serve a given embedding goal. For instance, to design an embedding method for repairing Rust programs, the designer might want to disregard certain elements, such as moving ownership to closure bodies. Such elements may be irrelevant for repair because they did not appear in the bug patterns that we presented.

All of our parsing, path extraction, weighting scheme, and clustering modules, along with our final results can be found in our replication package, which we provide for verification, reusability and further extension. Weightings can be modified to make the embedding focus on a specific set of elements for building different code embeddings.
